### 2013-07-29 appa.tex
where are my cannolis

% This is sec.tex
%
% This file is for appendices.  If you don't have any appendices, delete
% the percent sign from the beginning of the next line.
% \endinput

\appendix

% Add additional appendices by replicating the line below.

\section{An Extended Background}
\label{sec:prelim_background}
Let $W$ be a group with two generators $s$ and $t$ corresponding to reflecting over two hyperplanes in a real Euclidean space $V$, and let $\alpha_s$ and $\alpha_t$ be arbitrary vectors normal to these planes.  We will assume that the angle formed by $\alpha_s$ and $\alpha_t$ is not a rational multiple of $\pi$, so that the order of $st$ in $W$ is infinite.  In other words, $W$ is the infinite dihedral group with the presentation \[ W = \left<s,t \mid s^2=t^2=1\right>. \]

If $V$ has a positive definite symmetric bilinear form $(\mu, \lambda)$, then for any $a \in V$, reflection over the hyperplane normal to $a$ can be written explicitly for each $v \in V$ as the map \[ v \mapsto v - \frac{2(v,a)}{(a,a)} a. \]  
In particular, if $W$ acts on $V$ by this reflection, $s(\alpha_s) = -\alpha_s$ and $t(\alpha_t) = -\alpha_t$.  Furthermore, using the above, we see that there exist fixed constants $x$ and $y$ such that
\begin{align*}
	s(\alpha_t) &= \alpha_t + x \alpha_s \\
	t(\alpha_s) &= \alpha_s + y \alpha_t
\end{align*}
Explicitly, $x = \frac{-2(\alpha_s,\alpha_t)}{(\alpha_s, \alpha_s)}$ and $y = \frac{-2(\alpha_t,\alpha_s)}{(\alpha_t,\alpha_t)}$.

@@

We will consider maps between these bimodules that preserve left and right actions; that is, we consider the maps $\sigma$ such that
\begin{equation}
	\sigma(rx) = r\sigma(x) \quad\text{and}\quad \sigma(xr) = \sigma(x)r \quad\text{for every $r \in R$ and $x \in \text{Domain } \sigma$}.
	\label{eq:respect}
\end{equation}
These maps can be represented using diagrammatics, as below.

@@

Consider a category $\DD$ whose elements are linear combinations $\sum c_i \square$ of the diagrams described below (where the boxes are diagrams).  The coefficients $c_i$ belong to the ring $\ZZ[x,y]$.

@@

\subsection{Algebraic Context}
The maps are conventionally read from bottom to top.  The labels on the upper and lower boundaries specify the domain and codomain of the map by simply transcribing the tensor products to take -- for example, a map with bottom $stt$ and top $s$ represents a map from $R \stimes R \ttimes R \ttimes R$ to $R \stimes R$.  An unlabelled domain corresponds to $R$.

Let us hasten to introduce one final definition.
\begin{definition*}
	The \emph{Demazure operator} $\partial_s: R \to R$ is given by $\partial_s(f) = \frac{f - s(f)}{\alpha_s}$.
\end{definition*}

@@

Maps are composed by juxtaposition, and disjoint portions of the diagrams act independently.  Therefore, by a combination of these structures, we can generate arbitrarily complicated maps.  For example, the map in Figure~\ref{fig:example_diagram} represents a map \[ R \stimes R \ttimes R \stimes R \to R \stimes R \quad\text{by}\quad a \mid b \mid c \mid d \mapsto a \partial_s(bc) \mid d. \]

@@

\subsection{Operations for Diagrammatics}
\label{sec:prelim_genrel}
From the algebraic context given above, one can derive the following relationships, which are here encoded completely graphically.  These rules are sufficient to compute the maps we are interested in.

\setcounter{op}{-1}
\begin{op}[Isotropy] If two diagrams can be continuously deformed into each other (in the topological sense), then they are equivalent.  \end{op}
\begin{op}[Associativity] We have $\barbell{assoc_horiz} = \barbell{assoc_vert}$ and the similar equation for red.  \end{op}
\begin{op}[Contraction] We have $\barbell{contract_left} = \barbell{contract_right} = \barbell{alpha_blue}$ and the similar equation for red.  \end{op}
\begin{op}[The Needle] We have $\barbell{needle} = 0$.  Similarly, the red needle yields zero.  \end{op}
\begin{remark*} Using contraction and the above, one can show that the diagram $\barbell{zero}$ is zero as well. In fact, in general, \emph{any} map which contains a ``closed'' and empty region is the zero map.  \end{remark*}
\begin{op}[Barbell-Forcing Rules]
	We have the following three equalities:
	\begin{enumerate}[(a)]
		\ii $\barbell{barbell_blue}\barbell{alpha_blue} + \barbell{alpha_blue} \barbell{barbell_blue} = 2 \barbell{break_blue}$, and the similar equation for red.
		\ii $\barbell{alpha_red}\barbell{barbell_blue} = -x\barbell{break_red} + \barbell{barbell_blue}\barbell{alpha_red} + x \barbell{barbell_red}\barbell{alpha_red}$.
		\ii $\barbell{alpha_blue}\barbell{barbell_red} = -y\barbell{break_blue} + \barbell{barbell_red}\barbell{alpha_blue} + y \barbell{barbell_blue}\barbell{alpha_blue}$.
	\end{enumerate}
\end{op}

@@

Each bit of $\ul b$ is read from left to right in order and designated either up or down.  We say that a bit $x$ is an \emph{open neighbor} of $y$ if $x$ is marked up, $x$ and $y$ have the same color, and there are no other bits marked up between $x$ and $y$.
\begin{itemize}
	\ii If the bit is a $1$:
	\begin{itemize}
		\ii If the current bit has an open neighbor, then draw an arc joining the two bits, and mark both bits as down.
		\ii Otherwise, designate the current bit as up.
	\end{itemize}
	\ii If the bit is $0$:
	\begin{itemize}
		\ii If the current bit has an open neighbor, then draw an arc joining the two bits, mark the previous bit as down, and the current bit as up.
		\ii Otherwise, create a ``stub'', i.e. a single vertex of degree one, and mark the current bit as down.
	\end{itemize}
\end{itemize}
Finally, join any bits which are up at the end of the process to the upper boundary of the diagram.  A map which arises in such a manner is called a \emph{light leaf}.  An example of such a light leaf is shown in Figure~\ref{fig:lightleaf_example}.

@@

	Let us now compose $M_a \defeq \MM(\ul a)$ and $M_b \defeq \MM(\ul b)$ to obtain a full-map.  Two components (one from $M_a$ and one from $M_b$) are said to \emph{feel} each other $k$ times if they have $k$ antennae in common.  Notice that if any two components feel each other at least twice, then there exist two adjacent common antennae, whence the composition of the entire map is zero because the two antennae combine to form an empty bubble.  Furthermore, it is easy to see that any bubbles must be formed in this manner.  In combination with the claim above, this implies the first case of the theorem.

@@

\section{Proofs of Lemmata}
\label{sec:lemmata_proof}
\subsection{Proof of the Bubble Lemma}
We begin with a proof of the Bubble Lemma\footnote{I should remark that attempting to construct counterexamples to the Bubble Lemma leads to a far better understanding than trying to read the following proof.}, which is the crux result of the alternating case.  Indeed, both of the subsequent lemmas are fairly straightforward consequences of this one.

\begin{proof}[Proof of Lemma \ref{thm:bubble}]
	Take a bubble $B$ and assume without loss of generality that it is blue.  It is easy to verify that all caterpillars held inside $B$ are red, otherwise this contradicts the construction of Libedinsky's light leaves.  Furthermore, it is evident that the bubble is nonempty; hence there is at least one red caterpillar in $B$.  We will show that there is exactly one.

@@

	\begin{figure}[ht]
		\centering
		\begin{asy}
			size(6cm);
			pen s = blue, t = red + dashed + 0.6;
			pen dot_s = blue, dot_t = red;
			for (int i=0; i<=8; i+=2) { dot( (i,0), dot_s ); }
			for (int i=1; i<=8; i+=2) { dot( (i,0), dot_t ); }
			draw((0,0)..(1,2)..(3,4)..(5,3)..(6,5)..(7,2)..(8,0), s);
			draw((1,0)..(3,2)..(5,0), t); // red covering arc
			draw((2,0)..(3,1)..(4,0), s); // blue covering arc again
			label("$A$", (4.1,1.6), dir(60));
			label("$u$", (3,0), dir(-90));
			label("$x$", (4,0), dir(-90));
			label("$v$", (5,0), dir(-90));
		\end{asy}
		\hspace{2em}
		\begin{asy}
			size(6cm);
			pen s = blue, t = red + dashed + 0.6;
			pen dot_s = blue, dot_t = red;
			for (int i=0; i<=8; i+=2) { dot( (i,0), dot_s ); }
			for (int i=1; i<=8; i+=2) { dot( (i,0), dot_t ); }
			draw((0,0)..(1,2)..(3,4)..(5,3)..(6,5)..(7,2)..(8,0), s);
			draw((1,0)..(4,2)..(7,0), t); // red covering arc
			draw((2,0)..(4,1)..(6,0), s); // blue covering arc
			label("$A$", (5.6,1.5), dir(60));
			label("$u$", (3,0), dir(-90));
			label("$x$", (4,0), dir(-90));
			label("$v$", (5,0), dir(-90));
		\end{asy}
		\caption{Proving the bubble lemma with quasi-connected vertices.}
		\label{fig:pf_bubble}
	\end{figure}
	Because the string is alternating, we can take a blue anchor $x$ immediately to the left of $v$.  Now, there must be some blue arc shielding $u$ and $v$ (since $u,v$ are inside a blue bubble).  By the light-leave construction, there must be some red arc $A$ covering $x$.  If this arc joins $u$ and $v$, then we are done.  
	Otherwise, if it has endpoint $v$ then it must completely cover $u$ by condition (ii) of quasi-connectedness; if not, then it must completely cover $v$.  This leads us to two cases.
	\begin{itemize}
		\ii If the arc covers exactly one of $u$ or $v$, say $u$, then there must be a blue arc shielding $u$ from $A$.  But then this contradicts the quasi-connectedness, because now it is not possible to join $u$ and $v$ by an arc on this half-map.  This case is illustrated by the left half of Figure~\ref{fig:pf_bubble}.
		\ii If the arc covers both $u$ and $v$, then there must be a blue arc which shields both $u$ and $v$, which is itself shielded by $A$.  Then we can repeat the same procedure with this new blue arc.  This case is illustrated by the right half of Figure~\ref{fig:pf_bubble}.
	\end{itemize}
	Therefore any pair of quasi-connected anchors must be connected by an arc.  However, one can verify that the red anchors in $B$ but not in any other bubble are ``connected'' under quasi-connectedness.\footnote{That is, one can follow a sequence of red anchors, each consecutive pair quasi-connected, from any red anchor to any other red anchor.}  Therefore all such red anchors must be connected.  So there is a single caterpillar, as desired.
\end{proof}

@@

First, a few definitions for this section only.
\begin{definition*}
	Consider a caterpillar $C$ in a full-diagram with base $\AA_n$.  Define the \emph{nesting level} of the caterpillar as the largest integer $n$ for which there exists caterpillars $C = C_0$, $C_1$, $C_2$, $\dots$, $C_n$ such that $C_i$ holds $C_{i+1}$ for each $i=0,1,\dots,n-1$.  Define the \emph{size} of the caterpillar to be the number of bubbles it contains.
\end{definition*}

@@

	It is easy to verify that the top of this new map is zero.  Now what we have done is enclose the contents of each pasture into bubbles.  The Bubble Lemma tells us that the innermost bubble can have exactly one caterpillar, while the other bubbles must contain only a single caterpillar, which was once a fence.  This implies the Pasture Lemma in the first case.

>>>>

% This is appa.tex
\appendix

%TODO restate all theorems

\section{An Extended Background}
\label{sec:prelim_background}
Let $V$ be a (real) Euclidean space endowed with a positive definite symmetric bilinear form $(\mu, \lambda)$.
Let $W$ be a group with two generators $s$ and $t$ corresponding to reflecting over two hyperplanes in $V$, and let $\alpha_s$ and $\alpha_t$ be arbitrary vectors normal to these planes.  We will assume that the angle formed by $\alpha_s$ and $\alpha_t$ is not a rational multiple of $\pi$, so that the order of $st$ in $W$ is infinite.

In other words, $W$ is the infinite dihedral group with the presentation \[ W = \left<s,t \mid s^2=t^2=1\right>. \]

Then for any $a \in V$, reflection over the hyperplane normal to $a$ can be written explicitly for each $v \in V$ as the map \[ v \mapsto v - \frac{2(v,a)}{(a,a)} a. \]  
In particular, if $W$ acts on $V$ by this reflection, $s(\alpha_s) = -\alpha_s$ and $t(\alpha_t) = -\alpha_t$.  Furthermore, using the above, we see that there exist fixed constants $x$ and $y$ such that
\begin{align*}
	s(\alpha_t) &= \alpha_t + x \alpha_s \\
	t(\alpha_s) &= \alpha_s + y \alpha_t
\end{align*}
Explicitly, $x = \frac{-2(\alpha_s,\alpha_t)}{(\alpha_s, \alpha_s)}$ and $y = \frac{-2(\alpha_t,\alpha_s)}{(\alpha_t,\alpha_t)}$.

@@

We will consider maps between these bimodules that preserve left and right actions; that is, we consider the maps $\sigma$ such that
\begin{equation}
	\sigma(rx) = r\sigma(x) \quad\text{and}\quad \sigma(xr) = \sigma(x)r \quad\text{for every $r \in R$ and $x \in \text{Domain } \sigma$}.
	\label{eq:respect}
\end{equation}
These maps may be represented diagrammatically.

@@

Consider a category $\DD$ whose elements are linear combinations $\sum c_i \square$ of the diagrams described below (where the $\square$'s are diagrams).  The coefficients $c_i$ belong to the ring $\ZZ[x,y]$.

@@

\subsection{Algebraic Context}
The maps are conventionally read from bottom to top.  The labels on the upper and lower boundaries specify the domain and codomain of the map by ``transcribing'' the tensor products to take -- for example, a map with bottom $stt$ and top $s$ represents a map with domain $R \stimes R \ttimes R \ttimes R$ and codomain $R \stimes R$.  An unlabelled domain corresponds to $R$.

Let us introduce one final definition.
\begin{definition}
	The \emph{Demazure operator} $\partial_s: R \to R$ is given by \[ \partial_s(f) = \frac{f - s(f)}{\alpha_s}. \]
\end{definition}

@@

Maps are composed by juxtaposition, and disjoint portions of the diagrams act independently.  Therefore, by a combination of these structures, we can generate arbitrarily complicated maps.  For example, the diagram in Figure~\ref{fig:example_diagram} represents a map \[ R \stimes R \ttimes R \stimes R \to R \stimes R \quad\text{by}\quad a \mid b \mid c \mid d \mapsto a \partial_s(bc) \mid d. \]

@@

\subsection{Operations for Diagrammatics}
\label{sec:prelim_genrel}
From the algebraic context given above, one can derive the following relationships, which are here described completely graphically.  These rules are sufficient to compute the maps we are interested in.

Unless multiple colors appear in an equation, blue represents a generic color.

\setcounter{op}{-1}
\begin{op}[Isotropy] If two diagrams can be continuously deformed into each other (in the topological sense), then they are equivalent.  \end{op}
\begin{op}[Associativity] We have $\barbell{assoc_horiz} = \barbell{assoc_vert}$.  \end{op}
\begin{op}[Contraction] We have $\barbell{contract_left} = \barbell{contract_right} = \barbell{alpha_blue}$.  \end{op}
\begin{op}[The Needle] We have $\barbell{needle} = 0$.  \end{op}
\begin{remark*} Using contraction and the above, one can show that the diagram $\barbell{zero}$ is zero as well. In fact, in general, \emph{any} map which contains a ``closed'' and empty region is the zero map.  \end{remark*}
\begin{op}[Barbell-Forcing Rules]
	We have the following three equalities:
	\begin{enumerate}[(a)]
		\ii $\barbell{barbell_blue}\barbell{alpha_blue} + \barbell{alpha_blue} \barbell{barbell_blue} = 2 \barbell{break_blue}$, and the similar equation for red.
		\ii $\barbell{alpha_red}\barbell{barbell_blue} = -x\barbell{break_red} + \barbell{barbell_blue}\barbell{alpha_red} + x \barbell{barbell_red}\barbell{alpha_red}$.
		\ii $\barbell{alpha_blue}\barbell{barbell_red} = -y\barbell{break_blue} + \barbell{barbell_red}\barbell{alpha_blue} + y \barbell{barbell_blue}\barbell{alpha_blue}$.
	\end{enumerate}
\end{op}

@@

Each bit of $\ul b$ is read from left to right in order and designated either open or closed; the state of a bit can change more than once.  Initially all bits are closed.  We say that a bit $x$ is an \emph{open neighbor} of $y$ if $x$ is marked open, $x$ and $y$ have the same color, and there are no other bits marked open between $x$ and $y$.  Note that each bit will have at most one open neighbor.

\begin{itemize}
	\ii If the bit is a $1$:
	\begin{itemize}
		\ii If the current bit has an open neighbor, then draw an arc joining the two bits, and mark both bits as closed.
		\ii Otherwise, designate the current bit as open.
	\end{itemize}
	\ii If the bit is $0$:
	\begin{itemize}
		\ii If the current bit has an open neighbor, then draw an arc joining the two bits, mark the previous bit as closed, and the current bit as open.
		\ii Otherwise, create a ``stub'', i.e. a single vertex of degree one, and mark the current bit as down.
	\end{itemize}
\end{itemize}
Finally, join any bits which are open at the end of the process to the upper boundary of the diagram.  A map which arises in such a manner is called a \emph{light leaf}.  An example of such a light leaf is shown in Figure~\ref{fig:lightleaf_example}.

@@

	Let us now compose $M_a \defeq \MM(\ul a)$ and $M_b \defeq \MM(\ul b)$ to obtain a full-map.  Two components (one from $M_a$ and one from $M_b$) are said to \emph{feel} each other $k$ times if they have $k$ antennae in common.  Notice that if any two components feel each other at least twice, then there exist two adjacent common antennae, forcing the composition of the entire map to be zero because the two antennae combine to form an empty bubble.  Furthermore, it is evident that any bubbles must be formed in this manner.  In combination with the claim above, this implies the first case of the theorem.

@@

\section{Proofs of Lemmata}
\label{sec:lemmata_proof}
\subsection{Proof of the Bubble Lemma}
We begin with a proof of the Bubble Lemma\footnote{I should remark that attempting to construct counterexamples to the Bubble Lemma leads to a far better understanding than trying to read the following proof.}, which is the fundamental lemma for the alternating case.  Indeed, both of the subsequent lemmas are fairly straightforward consequences of this one.

\begin{proof}[Proof of Lemma \ref{thm:bubble}]
	Take a bubble $B$ and assume without loss of generality that it is blue.  By construction all caterpillars held inside $B$ must be red.  This implies that there must be at least one red caterpillar, because there is at least one red anchor inside $B$.

@@

	\begin{figure}[ht]
		\centering
		\begin{asy}
			size(6cm);
			pen s = blue, t = red + dashed + 0.6;
			pen dot_s = blue, dot_t = red;
			for (int i=0; i<=8; i+=2) { dot( (i,0), dot_s ); }
			for (int i=1; i<=8; i+=2) { dot( (i,0), dot_t ); }
			draw((0,0)..(1,2)..(3,4)..(5,3)..(6,5)..(7,2)..(8,0), s);
			draw((1,0)..(3,2)..(5,0), t); // red covering arc
			draw((2,0)..(3,1)..(4,0), s); // blue covering arc again
			label("$A$", (4.1,1.6), dir(60));
			label("$u$", (3,0), dir(-90));
			label("$x$", (4,0), dir(-90));
			label("$v$", (5,0), dir(-90));
		\end{asy}
		\hspace{2em}
		\begin{asy}
			size(6cm);
			pen s = blue, t = red + dashed + 0.6;
			pen dot_s = blue, dot_t = red;
			for (int i=0; i<=8; i+=2) { dot( (i,0), dot_s ); }
			for (int i=1; i<=8; i+=2) { dot( (i,0), dot_t ); }
			draw((0,0)..(1,2)..(3,4)..(5,3)..(6,5)..(7,2)..(8,0), s);
			draw((1,0)..(4,2)..(7,0), t); // red covering arc
			draw((2,0)..(4,1)..(6,0), s); // blue covering arc
			label("$A$", (5.6,1.5), dir(60));
			label("$u$", (3,0), dir(-90));
			label("$x$", (4,0), dir(-90));
			label("$v$", (5,0), dir(-90));
		\end{asy}
		\caption{Proving the bubble lemma with quasi-connected vertices.}
		\label{fig:pf_bubble}
	\end{figure}
	Because the string is alternating, we can take a blue anchor $x$ immediately to the left of $v$.  Now, there must be some blue arc shielding $u$ and $v$ (since $u,v$ are inside a blue bubble).  By construction, there must be some red arc $A$ covering $x$.  If this arc joins $u$ and $v$, then we are done.  
	Otherwise, if it has endpoint $v$ then it must completely cover $u$ by condition (ii) of quasi-connectedness; if not, then it must completely cover $v$.  This leads us to two cases.
	\begin{itemize}
		\ii If the arc covers exactly one of $u$ or $v$, say $u$, then there must be a blue arc shielding $u$ from $A$.  But then this contradicts the quasi-connectedness, because now it is not possible to join $u$ and $v$ by an arc on this half-map.  This case is illustrated by the left half of Figure~\ref{fig:pf_bubble}.
		\ii If the arc covers both $u$ and $v$, then there must be a blue arc which shields both $u$ and $v$, which is itself shielded by $A$.  Then we can repeat the same procedure with this new blue arc.  This case is illustrated by the right half of Figure~\ref{fig:pf_bubble}.
	\end{itemize}
	Therefore any pair of quasi-connected anchors must be connected by an arc.  However, one can verify that the red anchors in $B$ but not in any other bubble are ``connected'' under quasi-connectedness.\footnote{That is, one can follow a sequence of red anchors, each consecutive pair quasi-connected, from any red anchor to any other red anchor.}  Therefore all such red anchors must be connected.  So there is a single caterpillar, as desired.
\end{proof}

@@

First, a few definitions for this section only.
\begin{definition}
	Consider a caterpillar $C$ in a full-diagram with base $\AA_n$.  Define the \emph{nesting level} of the caterpillar as the largest integer $n$ for which there exists caterpillars $C = C_0$, $C_1$, $C_2$, $\dots$, $C_n$ such that $C_i$ holds $C_{i+1}$ for each $i=0,1,\dots,n-1$.  Define the \emph{size} of the caterpillar to be the number of bubbles it contains.
\end{definition}

@@

	Clearly the top of this new map is zero, say by \eqref{eq:top}.  Now what we have done is enclose the contents of each pasture into bubbles.  The Bubble Lemma tells us that the innermost bubble can have exactly one caterpillar, while the other bubbles must contain only a single caterpillar, which was once a fence.  This implies the Pasture Lemma in the first case.
